### 2012-03-13 dissertation-proposal/research-proposal.tex
Start on second draft

\documentclass[12pt, a4paper]{article}

\usepackage{biblatex}
\bibliography{bibliography}

\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\setlength{\topmargin}{-1.6cm}
\setlength{\leftmargin}{0.5cm}
\setlength{\rightmargin}{0.5cm}
\setlength{\textheight}{24.00cm} 
\setlength{\textwidth}{15.00cm}
\parindent 0pt
\parskip 5pt
\pagestyle{plain}

\title{Research Proposal}
\author{}
\date{}

\newcommand{\namelistlabel}[1]{\mbox{#1}\hfil}
\newenvironment{namelist}[1]{%1
\begin{list}{}
    {
        \let\makelabel\namelistlabel
        \settowidth{\labelwidth}{#1}
        \setlength{\leftmargin}{1.1\labelwidth}
    }
  }{%1
\end{list}}

\begin{document}
\maketitle

\begin{namelist}{xxxxxxxxxxxx}
\item[{\bf Title:}]
	Optional Static Type Checking for Dynamic Languages
\item[{\bf Author:}]
	Ambrose Bonnaire-Sergeant
\item[{\bf Supervisor:}]
	Rowan Davies
\end{namelist}

\section*{Background}

% In this section you should give some background to your
% research area. What is the problem you are tackling, and why is it
% worthwhile solving? Who has already done some work in this area,
% and what have they achieved?

Dynamic languages are designed to be convenient for the programmer to
write programs quickly and aspire to get out of the programmer's
way as much as possible. When programs grow large and stabilize,
the features of static languages are missed, specifically static
type checking.

Having a dynamic language that supports static properties 
has many potential practical benefits. 
Module-by-module porting of existing untyped code to a typed language 
is the largest benefit cited by Tobin-Hochstadt \cite{TypedScheme:2010}.
He developed Typed Scheme
\cite{TypedScheme:2010} 
to safely and incrementally port existing Scheme code, a dynamic language, to
Typed Scheme, a static language.

A recent example is Dart 
\cite{Dart:2012:Online}, a dynamically typed language that allows optional
static checks.

Good type systems deeply integrate with the language they are written for.
Statically typed languages usually include a type system in early releases,
which is able to evolve with the language while it is malleable.

The type system understands common idioms
and becomes the programmer's ally.

The tradeoffs of both styles has been subject to much debate, but some
points have general agreement. Dynamically typed languages
are generally more convenient for writing smaller 
programs, whereas statically typed languages are usually preferred 
when writing large programs.

Typed Racket \cite{TypedScheme:2010}
, which provides a statically typed variation of Racket 
with the intention of safely porting untyped code to typed code. 
Typed code can interact with untyped code while preserving some properties
of typed code by, for example, delaying static type checks until runtime.
Porting untyped Racket code to Typed Racket generally involves 
adding type annotations to top level declarations with minimal (usually zero)
modification to existing code structure.

Type checking is often an operation complected with compilation. This is
because languages that have type checkers generally enforce the rule that
if a program does not satisfy the type checker, then the program should not be 
allowed to compile. When considering an optional type checker, it makes sense
to provide type checking as a simple operation, callable a la carte. 
This relieves a common restriction associated with statically typed languages;
programs can now compile regardless of the results of the 
type checker.

\section*{Aim}

% Now state explicitly the hypothesis you aim to
% test. Make references to the items listed in the Reference section
% that back up your arguments for why this is a reasonable
% hypothesis to test, for example the work of Knuth~\cite{knuth}.
% Explain what you expect will be accomplished by undertaking this
% particular project.  Moreover, is it likely to have any other
% applications?

My goal is to develop an optional static type system for Clojure, 
intended for practical use, to be provided as a library.

It will be based on the lessons learnt throught the development
of Typed Scheme, and as a response to Tobin-Hochstadt's \cite{TypedScheme:2010}
suggestion to add type systems to dynamic languages.

I expect this library to be an important asset to Clojure's feature set.
I expect it will become a factor in language decision and justification for programming projects.

\section*{Method}

% In this section you should outline how you intend to go
% about accomplishing the aims you have set in the previous
% section. Try to break your grand aims down into small,
% achievable tasks. Try to estimate how long you will
% spend on each task, and draw up a timetable for each
% sub-task.

\begin{description}
\item[0.1]
	\begin{itemize}
	\item Concrete types (no type variables)
	\item Union types
	\item Fixed arity function types
	\item Typed deftype
	\item Uniform variable-arity function types
	\item Type checking without refinements or occurance typing
	\end{itemize}
\item[0.2]
	\begin{itemize}
	\item Occurance typing for type inference
	\item Type variables (polymorphic types)
	\end{itemize}
\item[0.3]
	\begin{itemize}
	\item Typed defrecord
	\item All Types annotated for clojure.core library
	\end{itemize}
\item[0.4]
	\begin{itemize}
	\item Manage interactions between typed namespaces
	\end{itemize}
\item[0.5]
	\begin{itemize}
	\end{itemize}
\item[0.6]
	\begin{itemize}
	\item Mutable reference types
	\item Non-uniform variable arity polymorphism
	\end{itemize}

\end{description}
\section*{Software and Hardware Requirements}
% Outline what your specific requirements will be with regard
% to software and hardware, but note that any special requests
% might need to be approved by your supervisor and the Head of
% Department.

Linux environment with Java, git, and maven installed.

% Overall, you should aim to produce roughly a two page document
% (and certainly no more than four pages)
% outlining your plan for the year.

% \begin{thebibliography}{9}
% \bibitem{knuth} D. E. Knuth. {\em The \TeX~book.}\/ Addison-Wesley,
% Reading, Massachusetts, 1984.
% \bibitem{lamport} L. Lamport. {\em \LaTeX~: A Document Preparation
% System}.\/ Addison-Wesley, Reading, Massachusetts, 1986.
% \bibitem{ken} Ken Wessen, Preparing a thesis using \LaTeX~, private
% communication, 1994.
% \bibitem{lamport2} L. Lamport. Document Production: Visual
% or Logical, {\em Notices of the Amer. Maths. Soc.},\/ Vol. 34,
% 1987, pp. 621-624.
% \end{thebibliography}

\printbibliography

\end{document}