### 2012-04-20 lit-review/lit-review-draft.tex
Add lit review draft

\documentclass[12pt, a4paper]{article}
\usepackage{biblatex}
\usepackage{fancyvrb}
\bibliography{bibliography}
\title{A Practical Optional Type System for Clojure, Literary Review}
\author{Ambrose Bonnaire-Sergeant (20350292)}

\begin{document}
\maketitle

\begin{abstract}
Clojure is a practical, dynamically typed language for the Java Virtual Machine.
I will design and implement a prototype optional static type 
system for Clojure, intended for practical use. Devising a static type system 
requires a good balance of components; they should be conceptually compatible with each other
while not significantly complicating its implementation.
I look at several novel inventions created as a result 
of the Typed Racket project.
Non-uniform variable-arity polymorphism and occurrence typing are two such
components that could be used in harmony in this project.
Typed Racket and Scala use different forms of local type inference,
which are compared and judged for applicability to this project.
Other relevant features of Scala are considered, such as its approach to bounded polymorphism,
the Scala class hierarchy, and the similarities of its methods of composition to Clojure's.
\end{abstract}

\tableofcontents

\section{Introduction}

Designing a static type system for any dynamically typed language requires
careful consideration. Dynamically typed languages usually have a particular
style, designed to be powerful yet non-intrusive. If a static type system
for such languages is to be effective, it should preserve existing idioms and
syntactic abstractions, at least until time tested idioms are devised for
the typed language \cite{SAMTH:dissertation} .

Typed Racket \cite{SAMTH:dissertation} integrates a selection of ideas to
create a typed sister language to Racket. 
Many ideas produced as a result of the Typed Racket project are relevant here.
Clojure and Racket are similar; they are both dynamically typed Lisps
and therefore have similar idioms. The inventions of
occurrence typing \cite{Tobin-Hochstadt:2010:LTU:1932681.1863561}
and a systematic approach to variable-arity polymorphism
\cite{Strickland:2009:PVP:1532974.1532978}
are both applicable to this project.

Clojure supports multimethods, a mechanism to dispatch on the
results of an arbitrary function. Statically typing multimethods that support
multiple dispatch (which are more specific than Clojure's multimethods) has been tackled
in the context of statically typed languages
\cite{Millstein02modularstatically}
. Typing multimethods that support dispatch on an arbitrary function
has not been approached in the context of a dynamic language utilizing 
occurrence typing.

Clojure and Scala \cite{Odersky06anoverview},
are also similar langages.  
They could both be considered object-oriented, being built on abstractions
and classes. While there are some significant differences between them,
Scala being statically typed for example, there are still ideas that are applicable to this project.

Scala's inclusion of bounded type variables are useful in the context
of a language that makes heavy use of subtyping. There are some parallels
between Clojure protocols, which is Clojure's main abstraction, and Scala
traits, both can conceptually be useful in a language with bounded polymorphism.

Scala's class hierarchy \cite{Odersky06anoverview}
is also interesting and relevant. Scala and Clojure are
both languages with a focus on host interoperability, and both run
on the Java Virtual Machine.  Scala's and Clojure's
differing treatment of Java's \emph{null} reference is
also worth reviewing when devising the class hierarchy for this project.

\section{Variable-Arity Polymorphism}

Functions with variable-arity are abundant in dynamically typed languages;
it is common to have a large standard library of functions that utilize
variable-arity, and Clojure is no different. 
Statically typed languages tend to have limited support
for typing variable-arity functions. Strickland claims 
\cite{Strickland:2009:PVP:1532974.1532978}
``no typed language supports them in a systematic and principled manner''.

Strickland et al.  \cite{Strickland:2009:PVP:1532974.1532978} distinguish between two kinds
of variable-arity functions: uniform and non-uniform.

Uniform variable-arity functions have a homogeneous variable parameter list.
For example, the Clojure \emph{+} function takes any number of numeric values
and returns their sum.
Statically typed languages such as Java and C\# provide support to type such functions,
since access to variable arguments occurs via a homogeneous array \cite{Strickland:2009:PVP:1532974.1532978}
.

Non-uniform variable-arity functions have a heterogeneous variable parameter list.
For example, Clojure's \emph{map} function takes a function and one or more sequences,
and applies the function to each element of the sequences simultaneously.
Languages like Java and C\# lack support for typing these kinds of functions,
neither do languages with more powerful type systems such as Haskell and ML.
Strickland et al. conjecture that ``if their type systems provided variable-arity
polymorphism, Haskell and ML programmers would routinely define such functions too''
\cite{Strickland:2009:PVP:1532974.1532978}
.

\emph{map} is also interesting because it treats polymorphic functions as first
class values. This is the key innovation of Strickland et al. . Dzeng and Haynes
\cite{Dzeng94typereconstruction} work on adding variable-arity functions to
an ML style language differs mostly in this respect. Dzeng and Hayes support uniform and
non-uniform variable-arity functions, but not first class polymorphic values.  \footnotemark

\footnotetext{
First class polymorphic functions are significant in the context of local type inference (next section).
Hosoya and Pierce \cite{Hosoya99howgood} identify that synthesizing the type
of a first class polymorphic function passed to a polymorphic function is
\emph{hard to synthesize}. This is relevant because, even though
Strickland et al. provide the technology to type non-uniform variable-arity
functions with first class polymorphic values, extra type annotations could still be
required to satisfy local type inference when passing a polymorphic function
as an argument. Odersky's \cite{Odersky:2001:CLT:373243.360207} local type inference algorithm
makes manually providing these annotations less frequent.
}

A common usage of variable parameters in Clojure (and Racket) is keyword
parameters. The paper \cite {Strickland:2009:PVP:1532974.1532978}
mentions keyword parameters briefly, but it is unclear how it fits
with the given calculus. Starred pre-types and dotted pre-types are used
as placeholders for uniform and non-uniform variable parameters respectively, but
keyword parameters tend to be grouped in two. They could be regarded as ``semi-uniform''
variable parameter list. The paper does not mention such a middle-ground 
between the two styles.

\section{Local Type Inference}

Local type inference is used in both Typed Racket 
\cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
and Scala \cite{Odersky06anoverview}, although both use slightly different
styles of checking. Typed Racket implements Local Type Inference
\cite{Pierce:2000:LTI:345099.345100},
while Scala implements Colored Local Type Inference \cite{Odersky:2001:CLT:373243.360207}.

Both local type inference papers describe similar algorithms. At the core
of any local type inference scheme is a \emph{bidirectional checking}
algorithm. This algorithm enables the local propagation of types both
up and down the syntax tree. Essentially, the algorithm is split into two
halves: checking mode and synthesis mode. In checking mode we are verifying
that an expression is a subtype of a given type, and in synthesis mode, the
types of expressions are propagated up when possible.

Colored Local Type Inference subsumes Local Type Inference. Colored Local Type Inference
allows the propagation of partial type information down the syntax tree,

Hosoya and Pierce \cite{Hosoya99howgood} identified two known issues with
local type inference, both related to synthesizing the type of arguments.
Unfortunately, they are unable to devise satisfactory extensions to existing
algorithms to mitigate these issues, and instead suggest avoiding
\emph{hard to synthesize} arguments. On the positive side, they identify some
common patterns for users to look out for. Odersky's \cite{Odersky:2001:CLT:373243.360207} 
algorithm is better behaved
than Piece and Turner's \cite{Pierce:2000:LTI:345099.345100} in some cases,
and is able to synthesize more \emph{hard to synthesize} expressions.

\section{Bounded and Unbounded Polymorphism}

The Local Type Inference paper \cite{Pierce:2000:LTI:345099.345100}
describes two implementations of type variables, for bounded
and unbounded type variables. The bounded implementation is presented
as an optional extension  to the unbounded implementation, which preserves all
properties described in the Local Type Inference algorithm.

An unbounded type variable does not have subtype constraints.
Bounded type variables can have subtype constraints, and 
subsume unbounded type variables \cite{Pierce:2000:LTI:345099.345100}
, as a unbounded variable can be represented as a variable bounded
by the Top type.

Still, unbounded type variables have an advantage: their implementations are
are simpler in the presence of a \emph{Bottom} type. 
The constraint resolution algorithm for bounded variables
is more subtle, due to ``some surprising interactions between bounded quantifiers
and the \emph{Bot} type'' \cite{Pierce:2000:LTI:345099.345100}, described fully
by Pierce \cite{Pierce97boundedquantification}.

Typed Racket \cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
supports unbounded polymorphism, while Scala \cite{Odersky06anoverview}
supports an extended form of bounded polymorphism called
F-bounded polymorphism \cite{Canning:1989:FPO:99370.99392}, which allows the
bound variable to occur in its own bound.
F-bounded polymorphism is useful in the context of object-oriented abstractions,
as demonstrated by Odersky \cite{Odersky06anoverview} .
This is one possible explanation why Typed Racket, which is not built on abstractions like Scala,
does not support bounded quantification. Unfortunately, no Typed Racket paper mentions 
bounded quantification, so the rationale is not clear.

Clojure, like Scala, is built on object-oriented abstractions. Clojure protocols
and Java interfaces (interfaces are supported by Clojure) are good candidates
for bounds in bounded or F-bounded polymorphism.

\section{Typed Racket}

Typed Racket is a statically typed sister language of Racket. It
attempts to preserve existing Racket idioms and aims type check
existing Racket code by simply adding top level type annotations. \cite{SAMTH:dissertation}

Typing macros is an open research problem \cite{Herman10:Theory},
Typed Racket expands all macro calls before type checking, \cite{SAMTH:dissertation}
avoiding the complex semantics of pre-expanded macro calls described by Herman 
\cite{Herman10:Theory} .
My design of the type system for Clojure will follow a similar strategy. Macros
will be expanded (by the compiler), and the expanded form will be type checked in
the normal fashion. Type checking macro definitions are outside the scope of this project.

Certain macros encode invariants that are not easily extracted. For example,
\emph{define-struct} in Racket defines a structure. Typed Racket includes the
\emph{define-struct:} macro \cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
, which annotates the structure's slots with types,
supports recursive types, and annotates the structure's type predicate with propositions
for compatibility with occurrence typing \cite{Strickland:2009:PVP:1532974.1532978}.
In this project, Clojure macros such as \emph{deftype} and \emph{defrecord}, used for defining a type
and a record respectively, would need similar typed versions.

Along with a full static type system, Typed Racket also comes with extensive support for
contracts to protect program invariants at runtime, \cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
building on Racket's existing infrastructure for contracts.
Adding contracts is outside the scope of this project, but would be considered desirable
future work.

Two other Typed Racket features worth mentioning are recursive types and refinement types  
\cite{SAMTH:dissertation}
.
Recursive types allow a type definition to refer to itself, enabling structurally
recursive types like binary trees. Refinement types let the programmer define
new types that are subsets of an existing type, such as the type for all even integer,
which is a subset of all integers.
Both these features would fit well in a future implementation of this project.

\section{Occurrence Typing}
\label{sec:OccurrenceTyping}

Dynamically typed languages use a combination of type predicates and
selectors to reason about the type of variables at various points in a program.

For example, the following snippet shows how the type predicate
\emph{number?} and the selector \emph{first} are used to direct control flow and ensure the 
type of local variables.

\newpage

\begin{Verbatim}[frame=single]
(let [x (list (number-or-string))]
  (cond 
    (number? (first x)) (+ 1 (first x))
    :else               (str (first x))))
\end{Verbatim}

If a static type system for a dynamically typed language 
should be made to support existing programming idioms, it must
be able to use the results of these predicates and selectors to
infer the types at particular points in the program.

Occurrence typing \cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
\cite{Tobin-Hochstadt:2010:LTU:1932681.1863561}
was developed for the Typed Racket project capture common Racket idioms.
Occurrence typing extends the type system with a \emph{proposition environment}
that represents the type refinements the type system has inferred
up to a point in the program.

The 2010 paper \cite{Tobin-Hochstadt:2010:LTU:1932681.1863561}
reformulated occurrence typing, improving the original 2008 paper
\cite{Tobin-Hochstadt:2008:DIT:1328897.1328486}
after revealing that ``three years of practical experience has revealed
serious shortcoming of our type system.'' \cite{Tobin-Hochstadt:2010:LTU:1932681.1863561}
The new implementation utilizes a simple proof system to solve
propositional logic statements, in terms of type predicates and selectors.

Type predicates and selectors require two extra annotations 
for the occurrence typing system to understand it: a ``then'' proposition
when the result is a true value the, and an ``else'' proposition for a false value.
For example, \emph{number?} would have a ``then'' proposition saying its argument
is a number, and an ``else'' proposition saying its argument is \emph{not} a number.

One extension not mentioned in the papers is tracking values of
key-value pairs in hash-maps. Tobin-Hochstadt \cite{Tobin-Hochstadt:2010:LTU:1932681.1863561}
uses the Racket selectors \emph{car} and \emph{cdr} to refine the type of
a list, but Clojure, which uses maps extensively, could benefit from refining the types 
of selectors like \emph{get} and \emph{find}, which extract values from maps.
This would require more fine grained types for hash-maps.

System D, a novel type system developed by Chugh et al. \cite{Chugh:2012:NRL:2103621.2103686}
, provides such fine grained hash-map types. By using \emph{nested refinements},
``System D can express complex invariants between base values and richer values''
\cite{Chugh:2012:NRL:2103621.2103686}
.
System D is much more expressive than occurrence typing, subsuming it, but it comes at a cost:
``the huge leap in expressiveness \ldots is accompanied by some unique technical challenges''
\cite{Chugh:2012:NRL:2103621.2103686}
.
An occurrence typing implementation is much more straightforward than System D, which involves
specially optimizing SMT queries to be fed into an SMT solver. By comparison, simple 
algorithms dealing with propositional logic are needed for occurrence typing.

Occurrence typing is a relatively simple, time worn technique used successfully 
in Typed Racket. Clojure is similar enough to Racket for occurrence typing to work
without issues, but further research should be carried out on more fine grained types
for hash maps to be satisfying pairing.

\section{Statically Typed Multimethods}

% challenges: methods occurring in different modules

A Clojure multimethod is slightly different to most other
abstractions of the same name. Here, a multimethod 
dispatches on the result of an \emph{arbitrary function},
rather that more traditional multimethods that dispatch on
the types or structure of its arguments.

Millstein and Chambers \cite{Millstein02modularstatically}
describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. They tackle a key challenge for statically typing
multimethods: ``it is possible for two modules containing arbitrary multimethods to typecheck
successfully in isolation but generate type errors when linked together.'' \cite{Millstein02modularstatically}

This challenge holds when statically typing Clojure multimethods.
One additional challenge would be multimethods being imported from 
untyped modules. This would have to be handled carefully.

If we consider a multimethod as a spread out conditional, a
novel use of occurrence typing becomes apparent. We can use
occurrence typing to refine the argument types in the method bodies
by utilizing the results of the dispatch function. This is simple
when dispatching by type, but a common idiom in Clojure is to dispatch
on the value associated with a key in a hash-map.

To satisfy this idiom, we require occurrence typing with more
finely grained hash-map types (see section~\ref{sec:OccurrenceTyping} for discussion).

\section{Conclusion}

Many related components must come together in the design of a
static type system. Typed Racket achieves a satisfying balance of 
occurrence typing, local type inference and variable-arity polymorphism.
Similarly, Scala features F-bounded polymorphism, a class hierarchy
that is compatible with Java, and colored local type inference.

I intend to design a static type system for Clojure
that features a satisfying combination of features, individually taken from other
languages, when combined produce novel and interesting problems.

I have outlined an interesting possible extension to occurrence typing,
adding finer grain hash-map types.

I have proposed novel method of statically typing multimethods
using occurrence typing to consider the results of the dispatch function.

I suggested combining these two ideas to produce an even more expressive
method for typing multimethods that dispatch on the value associated with keys in hash-maps.

\printbibliography

\end{document}