### 2012-09-11 dissertation/implementation.tex

\chapter{Implementation}

\section{Composite Propositions}

\section{Non-hygienic Macro Expansion and Filters}
% need to remove bindings as they fall out of scope
% Need hygiene to ensure correct propagation of filters
%(fn [a]
%  (if (= a 1)
%    (let [a 'foo] ; here this shadows the argument, impossible to recover filters
%      a)          ; in fact any new filters about a will be incorrectly assumed to be the argument
%      false))