### 2012-10-02 dissertation/experiments.tex

\chapter{Experiments}

% 
In this chapter we look at several example of using Typed Clojure and
how well the current prototype handles them.

\section{Java Interoperability}

As Typed Clojure is intended to be useful for practical purposes, it was important
to port existing code that utilized Java interoperability.
My porting effort attempted to follow how a real programmer might port code to Typed Clojure
and I describe my approach for each situation.

I ported a function from clojure.contrib.reflect\footnote{https://github.com/cemerick/pomegranate}, 
a Clojure library relying heavily on Java interoperability.
I chose \lstinline|call-method| for several reasons: it chains several Java calls together,
it uses primitive arrays, it exposes a wart of Clojure with \emph{Named} things,
and \lstinline|null| is a valid value in one place.

Before showing the implementation, there is a brief explanation of the relevant syntax.
Java methods are called using the \emph{dot} operator. If \lstinline|o| is an object \lstinline|(. o m a*)|
calls its method named \lstinline|m| passing \lstinline|a*| arguments. The method can be named first
using the equivalent sugar \lstinline|(.m o a*)|. Also, \lstinline|doto| is convenient 
notation for muliple method calls on the same object, presumably for side effects, and returns the original
object. For example, \lstinline|(doto o (.m1 a*) (m2 a*))| calls methods \lstinline|m1| and \lstinline|m2|
on \lstinline|o| and returns \lstinline|o|.

\begin{lstlisting}[caption=call-method, label=lst:callmethod]
(defn call-method
  "Calls a private or protected method.

  params is a vector of classes which correspond to the arguments to
  the method e

  obj is nil for static methods, the instance object otherwise.

  The method-name is given a symbol or a keyword (something Named)."
  [^Class klass method-name params obj & args]
  (let [method (doto (.getDeclaredMethod klass 
                                         (name method-name)
                                         (into-array Class params))
                 (.setAccessible true))]
    (.invoke method obj (into-array Object args))))
\end{lstlisting}

The original function is modified slightly for readability and is presented in listing \ref{lst:callmethod}.

\begin{lstlisting}[caption=call-method Type Annotation, label=lst:callmethodtype]
(ann call-method 
     [Class Named (IPersistentVector Class) (U nil Object) (U nil Object) * -> (U nil Object)])
\end{lstlisting}

Thankfully this function has up-to-date documention, and from it we can derive an expected type
(listing \ref{lst:callmethodtype}).

Before running the type checker we must convert our array constructors into ones that Typed Clojure
can understand. Array types in Typed Clojure are represented by \lstinline|(Array c t)|, where
the Java class \lstinline|c| is the Java component type,
and the Typed Clojure type \lstinline|t| is the Typed Clojure component type. 
We can pass this array to Java methods accepting
type \lstinline|c[]|, and we can read and write type \lstinline|t| to the array from Typed Clojure.

\begin{itemize}
\item \lstinline|(into-array Class params)| becomes \lstinline|(into-array> Class Class params)|
      which is of type \lstinline|(Array Class Class)|.
\item \lstinline|(into-array Object args)| becomes \lstinline|(into-array> Object (U nil Object) args)|
      which is of type \lstinline|(Array Object (U nil Object))|.
\end{itemize}

The second place in the \lstinline|Array| type constructor allows fine grained control over what is allowed
in the array. The first point above must be type \lstinline|(Array Class Class)| because the 
\lstinline|getDeclaredMethod| method on \lstinline|java.lang.Class| instances requires an array of non-null
\lstinline|Class| objects. On the other hand, the \lstinline|invoke| method takes an array that allows
\emph{null} members, so its type is \lstinline|(Array Object (U nil Object))|.

\section{Red-Black Tree}