### 2012-10-10 dissertation/experiments.tex
Monad experiment + other changes

\chapter{Experiments}

In this chapter we look at several example of using Typed Clojure and
how well the current prototype handles them. 
We intentially chose examples that could be challenging to type check
or were particularly useful to the everyday Clojure programmer.

\section{Java Interoperability}

As Typed Clojure is intended to be useful for practical purposes, it was important
to port existing code that utilized Java interoperability.
My porting effort attempted to follow how a real programmer might port code to Typed Clojure
and I describe my approach for each situation.

I ported a function from clojure.contrib.reflect\footnote{https://github.com/cemerick/pomegranate}, 
a Clojure library relying heavily on Java interoperability.
I chose \lstinline|call-method| for several reasons: it chains several Java calls together,
it uses primitive arrays, it exposes a wart of Clojure with \emph{Named} things,
and \lstinline|null| is a valid value in one place.

Before showing the implementation, there is a brief explanation of the relevant syntax.
Java methods are called using the \emph{dot} operator. If \lstinline|o| is an object \lstinline|(. o m a*)|
calls its method named \lstinline|m| passing \lstinline|a*| arguments. The method can be named first
using the equivalent sugar \lstinline|(.m o a*)|. Also, \lstinline|doto| is convenient 
notation for muliple method calls on the same object, presumably for side effects, and returns the original
object. For example, \lstinline|(doto o (.m1 a*) (m2 a*))| calls methods \lstinline|m1| and \lstinline|m2|
on \lstinline|o| and returns \lstinline|o|.

\begin{lstlisting}[caption=call-method, label=lst:callmethod]
;; call-method pulled from clojure.contrib.reflect, (c) 2010 Stuart Halloway & Contributors
(defn call-method
  "Calls a private or protected method.

  params is a vector of classes which correspond to the arguments to
  the method e

  obj is nil for static methods, the instance object otherwise.

  The method-name is given a symbol or a keyword (something Named)."
  [^Class klass method-name params obj & args]
  (let [method (doto (.getDeclaredMethod klass 
                                         (name method-name)
                                         (into-array Class params))
                 (.setAccessible true))]
    (.invoke method obj (into-array Object args))))
\end{lstlisting}

The original function is modified slightly for readability and is presented in listing \ref{lst:callmethod}.

\begin{lstlisting}[caption=call-method Type Annotation, label=lst:callmethodtype]
(ann call-method 
     [Class Named (IPersistentVector Class) (U nil Object) (U nil Object) * -> (U nil Object)])
\end{lstlisting}

Thankfully this function has up-to-date documentation, and from it we can derive an expected type
(listing \ref{lst:callmethodtype}).

Before running the type checker we must convert our array constructors into ones that Typed Clojure
can understand. Array types in Typed Clojure are represented by \lstinline|(Array c t)|, where
the Java class \lstinline|c| is the Java component type,
and the Typed Clojure type \lstinline|t| is the Typed Clojure component type. 
We can pass this array to Java methods accepting
type \lstinline|c[]|, and we can read and write type \lstinline|t| to the array from Typed Clojure.

\begin{itemize}
\item \lstinline|(into-array Class params)| becomes \lstinline|(into-array> Class Class params)|
      which is of type \lstinline|(Array Class Class)|.
\item \lstinline|(into-array Object args)| becomes \lstinline|(into-array> Object (U nil Object) args)|
      which is of type \lstinline|(Array Object (U nil Object))|.
\end{itemize}

The second place in the \lstinline|Array| type constructor allows fine grained control over what is allowed
in the array. The first point above must be type \lstinline|(Array Class Class)| because the 
\lstinline|getDeclaredMethod| method on \lstinline|java.lang.Class| instances requires an array of non-null
\lstinline|Class| objects. On the other hand, the \lstinline|invoke| method takes an array that allows
\emph{null} members, so its type is \lstinline|(Array Object (U nil Object))|.

Now we run the type checker, which produces a type error.

\begin{lstlisting}
#<Exception java.lang.Exception: 29: Cannot call instance method java.lang.reflect.AccessibleObject/setAccessible on type (U nil java.lang.reflect.Method)>
\end{lstlisting}

Because Typed Clojure assumes all methods return nilable Objects, the call to \lstinline|getDeclaredMethod|
has return type \lstinline|(U nil java.lang.reflect.Method)|. It is not type safe to call \lstinline|setAccessible|
on this type, so we get a type error.

In this case, Typed Clojure is too conservative: according to its documentation \lstinline|getDeclaredMethod|
never returns \emph{null}. We add this rule with \lstinline|non-nil-return|.

\begin{lstlisting}
(non-nil-return java.lang.Class/getDeclaredMethod :all)
\end{lstlisting}

Running the type checker produces a different type error.

\begin{lstlisting}
#<Exception java.lang.Exception: Type Error, REPL:32 - (U java.lang.Object nil) is not a subtype of: java.lang.Object>
\end{lstlisting}

This concerns passing \lstinline|obj| as the first argument to the \lstinline|invoke| method.
Typed Clojure conservatively defaults method parameter types as non-nullable. 
Therefore the first parameter of \lstinline|invoke| is \lstinline|Object| by default; \lstinline|obj|
is \lstinline|(U java.lang.Object nil)|. Again, this is too conservative as the first argument can
be \emph{null} for static methods, and we use \lstinline|nilable-param| to specify the first argument
of \lstinline|invoke| may be nil, for the arity of two parameters.

\begin{lstlisting}
(nilable-param java.lang.reflect.Method/invoke {2 #{0}})
\end{lstlisting}

Here is the final successfully type checked code.
\begin{verbatim}Is this a reasonable result?\end{verbatim}

\begin{lstlisting}
(non-nil-return java.lang.Class/getDeclaredMethod :all)
(nilable-param java.lang.reflect.Method/invoke {2 #{0}})

(ann call-method [Class Named (IPersistentVector Class) (U nil Object) (U nil Object) * -> (U nil Object)])

;; call-method pulled from clojure.contrib.reflect, (c) 2010 Stuart Halloway & Contributors
(defn call-method
  "Calls a private or protected method.

  params is a vector of classes which correspond to the arguments to
  the method e

  obj is nil for static methods, the instance object otherwise.

  The method-name is given a symbol or a keyword (something Named)."
  [^Class klass method-name params obj & args]
  (let [method (doto (.getDeclaredMethod klass 
                                         (name method-name)
                                         (into-array> Class Class params))
                 (.setAccessible true))]
    (.invoke method obj (into-array> Object (U nil Object) args))))
\end{lstlisting}

\section{Red-Black Tree}

\section{Monads}

Monads[] are an interesting control structure probably most recognisable from its inclusion
in the statically-typed language Haskell[].

I chose to port a subset of the Clojure library \emph{algo.monads} to Typed Clojure. The library
provides support for several kinds of monads.

This library represents a monad as a hash-map with four keys: \lstinline|:m-bind|, \lstinline|:m-result|,
\lstinline|:m-zero|, and \lstinline|:m-plus|. A valid monad must provide the first two, and the latter
two may optionally be mapped to the keyword \lstinline|::undefined|\footnote{Keywords prefixed with \lstinline|::|
are qualified in the current namespace.}.

\begin{lstlisting}[caption=Untyped definition for the identity monad, label=lst:identitymdef]
(defmonad identity-m
   "Monad describing plain computations. This monad does in fact nothing
    at all. It is useful for testing, for combination with monad
    transformers, and for code that is parameterized with a monad."
  [m-result identity
   m-bind   (fn m-result-id [mv f]
              (f mv))
  ])
\end{lstlisting}

\begin{lstlisting}[caption=Macro expansion of identity monad definition, label=lst:identitymexpand]
(def identity-m
  (let [m-bind :typed.test.monads/undefined
        m-result :typed.test.monads/undefined
        m-zero :typed.test.monads/undefined
        m-plus :typed.test.monads/undefined
        m-result identity
        m-bind (fn m-result-id [mv f] (f mv))]
    {:m-plus m-plus, :m-zero m-zero, :m-bind m-bind, :m-result m-result}))
\end{lstlisting}

A monad is defined using the macro \lstinline|defmonad|. The macro expands to code that binds a var to a hash-map
with the aforementioned keys. For example, listing \ref{lst:identitymdef} expands to 
the code in listing \ref{lst:identitymexpand}.

\begin{lstlisting}[caption=Type for identity monad, label=lst:identitymtype]
(ann identity-m
     '{:m-bind (All [x y]
                 [x [x -> y] -> y])
       :m-result (All [x]
                   [x -> x])
       :m-zero '::undefined
       :m-plus '::undefined})
\end{lstlisting}

Typed Clojure operates directly on fully macro-expanded forms, so it is unsurprising that the type
annotation for \lstinline|identity-m| reflects its macro-expansion (listing \ref{lst:identitymtype}).
The \lstinline|:m-bind| entry is comparable to Haskell's monadic \emph{bind},
and the \lstinline|:m-result| entry is comparable to Haskell's monadic \emph{return}.

\begin{lstlisting}[caption=Maybe monad definition, label=lst:maybemdef]
(defmonad maybe-m
  "Monad describing computations with possible failures. Failure is
  represented by nil, any other value is considered valid. As soon as
  a step returns nil, the whole computation will yield nil as well."
  [m-zero   nil
   m-result (fn m-result-maybe [v] 
              v)
   m-bind   (fn m-bind-maybe [mv f]
              (if (nil? mv) nil (f mv)))
   m-plus   (fn m-plus-maybe [& mvs]
              (first 
                (filter #(not (nil? a))) mvs))
   ])
\end{lstlisting}

Instead of explaining the identity monad further, we show a more interesting monad, the maybe monad.
Its untyped definition is given in listing \ref{lst:maybemdef}.

\begin{lstlisting}[caption=Typed maybe monad definition, label=lst:maybemtyped]
(ann maybe-m
    '{:m-bind (All [x y]
                [(U nil x) [x -> (U nil y)] -> (U nil y)])
      :m-result (All [x]
                  [x -> (U nil x)])
      :m-zero nil
      :m-plus (All [x]
                [(U nil x) * -> (U nil x)])})
(defmonad maybe-m
  "Monad describing computations with possible failures. Failure is
  represented by nil, any other value is considered valid. As soon as
  a step returns nil, the whole computation will yield nil as well."
  [m-zero   nil
   m-result (ann-form 
              (fn m-result-maybe [v] v)
              (All [x] 
                [x -> (U nil x)]))
   m-bind   (ann-form 
              (fn m-bind-maybe [mv f]
                (if (nil? mv) nil (f mv)))
              (All [x y]
                [(U nil x) [x -> x] -> (U nil x)]))
   m-plus   (ann-form 
              (fn m-plus-maybe [& mvs]
                (first 
                  (filter
                    (ann-form 
                      #(not (nil? %))
                      [(U nil x) -> boolean :filters {:then (is x 0)}])
                    mvs)))
              (All [x]
                [(U nil x) * -> (U nil x)]))
   ])
\end{lstlisting}

A significant difference between typing Haskell style monads in Haskell and typing monads in \emph{algo.monads} with Typed Clojure
is that monads in Haskell are represented by type classes, while Typed Clojure does not support higher-kinded
constructs.

\section{Conduit}