### 2012-10-10 dissertation/paper.tex

\documentclass{cshonours}

\usepackage{mmm}
\usepackage{mathpartir}
\usepackage{clj-grammar}
\usepackage[style=numeric]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{listings}
\lstset{ %
  language=Lisp,                % choose the language of the code
  columns=fixed,basewidth=.5em,
  basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
  %numbers=left,                   % where to put the line-numbers
  %numberstyle=\small\ttfamily,      % the size of the fonts that are used for the line-numbers
  %stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
  %numbersep=5pt,                  % how far the line-numbers are from the code
  %backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  %showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  %showtabs=false,                 % show tabs within strings adding particular underscores
  %frame=single,           % adds a frame around the code
  %tabsize=2,          % sets default tabsize to 2 spaces
  captionpos=t,           % sets the caption-position to bottom
  breaklines=true,        % sets automatic line breaking
  breakatwhitespace=true,    % sets if automatic breaks should only happen at whitespace
  %escapeinside={\%*}{*)},          % if you want to add a comment within your code
}

\title{A Practical Optional Type System for Clojure}
\author{Ambrose Bonnaire-Sergeant}

\keywords{Type systems, Clojure}
% TODO What are categories? cshonours requires them
\categories{}

\begin{document}

\maketitle

\begin{abstract}
  % FIXME long tradition for attempts
  Since the first wide-spread programming languages like Fortran, LISP and ALGOL
  there has been a division between languages that perform type checking at compile-time
  and those that don't.
  A recent trend is to combine the advantages of both kinds of languages by adding optional
  static type systems to languages without static type checking,
  which promises to eliminate many common user errors
  that are otherwise unnoticed or are caught when the program is run.

\begin{verbatim}
  Recent attempts have seen several promising and innovative
  type systems built for dynamically typed languages. Typed Racket is able to
  statically type check idioms ..
\end{verbatim}
  We take the lessons learnt from these
  projects and apply them to the dynamically-typed language Clojure in the form of Typed Clojure.
\end{abstract}

%\tableofcontents

\include{intro}
\include{lit-review}
\include{design-choices}

% A syntactic approach to type soundness (Wright and Felleisen)
% operational semantics
% - See Sam TH's operational semantics
% TAPL 392

% potential proofs/rules
% - hmaps
% - intersection types (protocols + types)
% - subtyping
% - dot methods
% - f-bounded polymorphism

\include{implementation}

% Evaluation
% - successes
% - future work identified

\include{experiments}

\include{future-work}

%
%\begin{mathpar}
%\inferrule*[right=B-Var]
%  {\rho(x) = v}
%  {\rho \vdash x \downarrow v}
%
%\inferrule*[right=B-Assoc]
%  {\rho \vdash e_m \downarrow mi \\
%   \rho \vdash e_k \downarrow k \\
%   \rho \vdash e_v \downarrow v \\
%   \rho \vdash \overrightarrow{e_j}^n \downarrow \overrightarrow{k_j}^n \\
%   \rho \vdash \overrightarrow{e_m}^n \downarrow \overrightarrow{k_m}^n \\
%   \delta(add\_entry, mi, k, v, \overrightarrow{k_j, k_m}^n) = mo}
%  {\rho \vdash (assoc\ e_m\ e_k\ e_v\ \overrightarrow{e_j\ e_m}^n) \downarrow mo}

%\infrule[B-Var]
%  {\rho(x) = v}
%  {\rho \vdash x \downarrow v}
%
%\infrule[B-Assoc]

%\end{mathpar}

\printbibliography[title=References]

\end{document}