### 2012-10-22 dissertation/future-work.tex
Example for infer

\chapter{Future Work}

\section{Multimethods}
\label{sec:mmfuture}

\begin{verbatim}
Multimethods in Clojure are an idiomatic and often used feature.
It would be essential for an optional static type system for
Clojure to support
There are at least two hurdles to supporting multimethods.
\end{verbatim}

% Problem - typing isa?
% - probably impossible to type higher order usages of isa?.
% Dispatch function must infer latent filters for multi-arity functions. (Typed Racket only infers single arg I think)
% Statically proving at least one relevant dispatch method is present.
% Warn potential method conflicts.
% Suppress warnings, similar to prefer-method syntax

\section{Records}
% A composition of existing features: Datatypes and Heterogeneous Maps

A Clojure record is a composite of datatypes and maps; it is a datatype
with fields that can be treated like a map (records support map operations like \lstinline|get|,
\lstinline|assoc|, and \lstinline|dissoc|).
Records have some quirks that may prove difficult to

\subsection{Other Clojure Dialects}
% unify ASTs for easier porting

\subsection{Linter}