### 2012-10-22 dissertation/future-work.tex

\chapter{Future Work}

\section{Contracts and Blame}
\label{sec:contractsblame}

\section{Multimethods}
\label{sec:mmfuture}

\begin{verbatim}
Multimethods in Clojure are an idiomatic and often used feature.
It would be essential for an optional static type system for
Clojure to support
There are at least two hurdles to supporting multimethods.
\end{verbatim}

% Problem - typing isa?
% - probably impossible to type higher order usages of isa?.
% Dispatch function must infer latent filters for multi-arity functions. (Typed Racket only infers single arg I think)
% Statically proving at least one relevant dispatch method is present.
% Warn potential method conflicts.
% Suppress warnings, similar to prefer-method syntax

\section{Records}

A Clojure record is a composite of datatypes and maps; it is a datatype
with fields that can be treated like a map (records support map operations like \lstinline|get| (lookup),
\lstinline|assoc| (add entries), and \lstinline|dissoc| (remove entries)).
One interesting property from the perspective of static typing is that if
a entry corresponding to a field of the underlying datatype is removed
with \lstinline|dissoc|, then the resulting type is a plain map: it loses
its record type.
Correspondingly, a record keeps its type if any other entry is removed.

Records are used frequently in Clojure code, so it is desirable
to support them to some degree. Future work is planned to investigate solutions
to satisfactorily capture their subtle semantics in a practical way.

\subsection{Other Clojure Dialects}
% unify ASTs for easier porting

\begin{verbatim}
Other Clojure implementations
\end{verbatim}

\subsection{Java Generics}

\begin{verbatim}
Currently, Typed Clojure does not interact with Java generics,
which provide
instead 
It may be desirable to 
\end{verbatim}

\subsection{Linter}