### 2012-10-24 dissertation/paper.tex

\documentclass{cshonours}

\usepackage{mmm}
\usepackage{mathpartir}
\usepackage{clj-grammar}
\usepackage{url}
\usepackage[style=alphabetic]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{listings}
\lstset{ %
  language=Lisp,                % choose the language of the code
  columns=fixed,basewidth=.5em,
  basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
  %numbers=left,                   % where to put the line-numbers
  %numberstyle=\small\ttfamily,      % the size of the fonts that are used for the line-numbers
  %stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
  %numbersep=5pt,                  % how far the line-numbers are from the code
  %backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  %showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  %showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,           % adds a frame around the code
  %tabsize=2,          % sets default tabsize to 2 spaces
  captionpos=t,           % sets the caption-position to bottom
  breaklines=true,        % sets automatic line breaking
  breakatwhitespace=true,    % sets if automatic breaks should only happen at whitespace
  %escapeinside={\%*}{*)},          % if you want to add a comment within your code
}

\title{A Practical Optional Type System for Clojure}
\author{Ambrose Bonnaire-Sergeant}

\keywords{Type systems, Clojure}
% TODO What are categories? cshonours requires them
\categories{}

\begin{document}

\maketitle

\begin{abstract}

Traditionally there has been a clear distinction between 
programming languages that perform type checking at compile-time, and those that do not.
In particular, the early wide-spread programming languages ALGOL and Lisp
could be divided this way: ALGOL performs compile-time (or static) type checking,
and Lisp does not, and most recent languages can be similarly classified.
Static type checking eliminates many common user errors that are otherwise unnoticed or are caught 
when the program is run.
On the other hand, a language without static type checking can often be designed 
to be more flexible and forgiving, making it easy, for example, to have a \emph{map} function that can map functions
with any number of arguments (as in most Lisp dialects).

A recent trend is to aim to combine the advantages of both kinds of languages by adding optional static 
type systems to languages without static type checking. 
A major challenge in these combinations is continuing to support the idiomatic style of the original 
language while performing accurate enough checking to catch similar errors to standard statically
typed languages. Some recent attempts appear to have finally produced type systems that have overcome this obstacle.
Notably Typed Racket, a typed sister language of the Lisp dialect Racket, is able to statically type 
check most idiomatic Racket code.
These optional type systems have some important differences from standard static
type systems, for example they often include \emph{union} types.

Clojure is a recent and increasingly popular programming language that is designed to support 
functional programming and immutability by default. Implementations of Clojure are hosted 
on popular platforms and provide direct interoperability with the host platform.
There are several implementations of Clojure, each targeting different runtimes including
the Java Virtual Machine, the Common Language Runtime, and Javascript virtual machines.

This dissertation describes my work on designing an optional static type system for Clojure based on
the lessons learnt from several projects, primarily Typed Racket.
This work includes designing and building a type checker for Clojure running on the Java Virtual Machine.
Several experiments are conducted using this prototype, particularly involving
existing Clojure code that is sufficiently complicated that
type checking increases confidence that the code is correct.
For example, nearly all of \emph{algo.monads}, a Clojure Contrib library for monadic programming, 
is able to be type checked.
Most monad, monad transformer, and monadic function definitions can be type checked,
usually by adding type annotations in natural places like function definitions.

There is significant future work to fully type check all Clojure features and idioms.
For example, multimethod definitions and functions accepting an even number of variable arguments
are troublesome. 
Also, there are desirable features from the Typed Racket project that are missing, such
as automatic runtime contract generation and a sophisticated blame system, 
both which are designed to improve error messages when mixing typed and untyped code in similar systems.
Overall, the work described in this dissertation leads to the conclusion that it appears to 
be both practical and useful to design and implement an optional static type system for the
Clojure programming language.

\end{abstract}

\tableofcontents

\include{intro}
\include{lit-review}
\include{design-choices}

% A syntactic approach to type soundness (Wright and Felleisen)
% operational semantics
% - See Sam TH's operational semantics
% TAPL 392

% potential proofs/rules
% - hmaps
% - intersection types (protocols + types)
% - subtyping
% - dot methods
% - f-bounded polymorphism

\include{implementation}

\include{experiments}

\include{future-work}

\chapter{Conclusion}

Whether a language is dynamically typed or statically typed is starting to
blur. Recent languages have managed to support the advantages of both styles:
the safety of a statically typed language while retaining the idioms found in
dynamically typed languages.

Extensions like Typed Clojure

\begin{verbatim}
  Programming is an error-prone task, and verification tools are important.
  Dynamic languages rely on several techniques for verification, but static
  type checking is not often supported.
\end{verbatim}

\printbibliography[title=References]

\include{appendix}

\end{document}