### 2012-12-06 dissertation/paper.tex
Add Supervisor. Revise abstract, motivation and acknowledgements

\documentclass{cshonours}

\usepackage{mmm}
\usepackage{mathpartir}
\usepackage{clj-grammar}
\usepackage{url}
\usepackage[style=alphabetic]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{listings}
\lstset{ %
  language=Lisp,                % choose the language of the code
  columns=fixed,basewidth=.5em,
  basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
  %numbers=left,                   % where to put the line-numbers
  %numberstyle=\small\ttfamily,      % the size of the fonts that are used for the line-numbers
  %stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
  %numbersep=5pt,                  % how far the line-numbers are from the code
  %backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  %showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  %showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,           % adds a frame around the code
  %tabsize=2,          % sets default tabsize to 2 spaces
  captionpos=t,           % sets the caption-position to bottom
  breaklines=true,        % sets automatic line breaking
  breakatwhitespace=true,    % sets if automatic breaks should only happen at whitespace
  %escapeinside={\%*}{*)},          % if you want to add a comment within your code
}

\title{A Practical Optional Type System for Clojure}
\author{Ambrose Bonnaire-Sergeant}

\keywords{Programming languages, optional type systems, Clojure}
% TODO What are categories? cshonours requires them - but commented out now
%\categories{}

\begin{document}

\maketitle

\begin{abstract}

%Traditionally there has been a clear distinction between 
%programming languages that perform type checking at compile-time, and those that do not.
%In particular, the early wide-spread programming languages ALGOL and Lisp
%could be divided this way: ALGOL performs compile-time (or static) type checking,
%and Lisp does not, and most recent languages can be similarly classified.
%Static type checking eliminates many common user errors that are otherwise unnoticed or are caught 
%when the program is run.
%On the other hand, a language without static type checking can often be designed 
%to be more flexible and forgiving, making it easy, for example, to have a \emph{map} function that can map functions
%with any number of arguments (as in most Lisp dialects).
%
%A recent trend is to aim to combine the advantages of both kinds of languages by adding optional static 
%type systems to languages without static type checking. 
%A major challenge in these combinations is continuing to support the idiomatic style of the original 
%language while performing accurate enough checking to catch similar errors to standard statically
%typed languages. Some recent attempts appear to have finally produced type systems that have overcome this obstacle.
%Notably Typed Racket, a typed sister language of the Lisp dialect Racket, is able to statically type 
%check most idiomatic Racket code.
%These optional type systems have some important differences from standard static
%type systems, for example they often include \emph{union} types.
%

Dynamic programming languages often abandon the advantages of static type checking 
in favour of their characteristic convenience and flexibility.
Static type checking eliminates many common user errors at compile-time that
are otherwise unnoticed, or are caught later in languages without static type checking.
A recent trend is to aim to combine the advantages of both kinds of languages by adding \emph{optional} static 
type systems to languages without static type checking, while preserving the languages idioms and style.

This dissertation describes my work on designing an optional static type system for the Clojure programming language,
a dynamically typed dialect of Lisp, based on
the lessons learnt from several projects, primarily Typed Racket.
This work includes designing and building a type checker for Clojure running on the Java Virtual Machine.
Several experiments are conducted using this prototype, particularly involving
existing Clojure code that is sufficiently complicated that
type checking increases confidence that the code is correct.
For example, nearly all of \emph{algo.monads}, a Clojure Contrib library for monadic programming, 
is able to be type checked.
Most monad, monad transformer, and monadic function definitions can be type checked,
usually by adding type annotations in natural places like function definitions.

There is significant future work to fully type check all Clojure features and idioms.
For example, multimethod definitions and functions with particular constraints on the number of 
variable arguments they accept (particularly functions taking only an even number of variable arguments) are troublesome. 
Also, there are desirable features from the Typed Racket project that are missing, such
as automatic runtime contract generation and a sophisticated blame system, 
both which are designed to improve error messages when mixing typed and untyped code in similar systems.

Overall, the work described in this dissertation leads to the conclusion that it appears to 
be both practical and useful to design and implement an optional static type system for the
Clojure programming language.

\end{abstract}

\begin{acknowledgements}

The past year has included by far my most enjoyable and motivating programming-related experiences.

Last November I attended my first programming conference, Clojure Conj 2011 (Raleigh, NC),
where I delivered a talk on logic programming with Clojure and met some of my programming heroes.
A week later my supervisor Rowan Davies invited me to visit Carnegie Mellon University
where he worked on his PhD project. Most memorably, I was invited to a meeting between
Rowan and his PhD supervisor Frank Pfenning.

This trip was funded by donations from the programming community. Thank you to all who helped me get there:
this work was directly inspired from conversations during this period.

Many thanks go to my supervisor Rowan. His influence as a lecturer inspired my exploration of
functional programming languages. He has exceeded my expectations as an honours supervisor,
and we plan to continue to collaborate in this area.
Rowan's experience with refinement types brought a unique perspective to this project
which was very welcome.

I am grateful to Sam Tobin-Hochstadt for his pioneering work in this area of research.
He has always been happy to offer assistance and advise throughout this project, which always proved invaluable.

I would also like to thank David Nolen for mentoring this project as part of Google Summer of Code 2012,
and Chas Emerick, without whom I would have never even considered attending Clojure Conj 2011
(and this work probably would not have happened).
Also, Daniel Spiewak, Dan Friedman and William Byrd, Jim Duey, Baishampayan Ghose, Sam Aaron, and
Steve Strickland, thank you for your help and inspiration.
Thank you Rich Hickey for creating Clojure, it is a thrill to build tools that make
this great language even better.

Fittingly, I gave a talk on Typed Clojure at Clojure Conj 2012, which was extremely satisfying.

Finally, thanks to my family, friends and to my girlfriend Tamara, each for your love and support.

\end{acknowledgements}

\tableofcontents

\include{intro}
\include{lit-review}
\include{design-choices}

% A syntactic approach to type soundness (Wright and Felleisen)
% operational semantics
% - See Sam TH's operational semantics
% TAPL 392

% potential proofs/rules
% - hmaps
% - intersection types (protocols + types)
% - subtyping
% - dot methods
% - f-bounded polymorphism

\include{implementation}

\include{experiments}

\include{future-work}

\chapter{Conclusion}

Whether a language is dynamically typed or statically typed is not always a
dividing classification. Recent languages have managed to support the advantages of both styles:
the safety of a statically typed language while retaining the idioms found in
dynamically typed languages.

This dissertation describes an optional static typing system
that attempts to bring the advantages of static typing to the dynamically typed language
Clojure, running on the Java Virtual Machine. 
By reusing many of the design choices made in similar projects
like Typed Racket, we are able to design and implement a prototype type system
\emph{Typed Clojure} that can statically check many Clojure idioms.

Typed Clojure is intended to be of practical use to Clojure programmers.
We show that Typed Clojure helps verify the absence of errors in Clojure code written in sophisticated
programming styles by porting most of a wide-spread Clojure library for monadic programming.
Similarly, we show that Typed Clojure helps verify code using Java interoperability is correct, and is
a useful tool to show misuse of Java's \emph{null}.

There is still significant future work in order to type check all Clojure idioms,
but the work already carried out suggests that an optional type system for Clojure
like Typed Clojure is both practical and useful tool.

\printbibliography[title=References]

\include{appendix}

\end{document}